A Quantitative Genetic Linked Selection Model fit to the Negative Selection Process in Humans

### 2023-07-06 manuscript/manuscript.tex
more ms stuff

However, across all of our annotation models, new deleterious mutations in
conserved feature classes were consistently estimated to have strongly
deleterious effects (Figure \ref{fig:figure-3}A), consistent with previous work
\parencite{McVicker2009-ax,Murphy2022-sj}. This includes features such as
coding sequences, PhastCons regions, or the top 6\% of basepairs with the most
pathogenic CADD scores. However, we find that there is considerable uncertainty
in how strong selection is against the most deleterious mutations. For example,
in the CADD 6\% full-track model, new mutations are estimated to have a
selection coefficients of $s = 10^{-1}$ with 42\% chance, $s=10^{-2}$ with
7.8\% chance, and $s=10^{-3}$ with 41\% chance. This creates a bimodal DFE
(Figure \ref{fig:figure-3}A, second row, left column), which is mirrored across
the other CADD models. Moreover, these DFE estimates suggest a very high
average selection coefficient across CADD 6\% regions of $\wisehat{\bar{s}}
\approx 0.038$.

Given that previous methods have typically not estimated the DFE up to
selection coefficients of $s=10^{-1}$, we ran additional ``constrained grid"
models using a DFE only up to $s=10^{-2}$. Across all populations, these
constrained grid models fit the data less well than the default grid. On
average, the constrained grid fits had $R_\text{LOCO}^2$ one percentage point
lower (see Supplementary Table \ref{supp:tbl-r2}). As expected, the constrained
grid models distributed mass onto the next strongest selection coefficient,
$s=10^{-2}$. While the DFE could be truly bimodal for highly-conserved regions
with a very high average selection coefficient, a likely alternative is that
these results reflect a non-identifiability issue between selective effects of
$s=10^{-1}$ and $s=10^{-2}$. We will refer to this as the
\emph{strong-selection non-identifiability} issue.

@@

Finally, we note that mutation rate estimates in the Feature Priority model are
unreasonably high ($\widehat{\mu} \approx 4.2 \times 10^{-8}$), reminiscent of
the high mutation rate estimates found under McVicker et al.'s
(\citeyear{McVicker2009-ax}) model. While both our and Murphy et al.'s CADD and
PhastCons-based models alleviate this issue, it is worth discussing why this
could occur. We gain some insight from comparing the estimated mutation rates
between our Feature and PhastCons Priority annotation models, which both
contain the exact same number of basepairs, but these are assigned to different
feature classes based on the priority of overlapping features. That one of
these models is our best-fitting model, and the other our worst indicates a
large role in per-basepair heterogeneity in selective effects in determining
the genome-wide linked selection signal. CADD-based models fit well in part due
to their fine-scale resolution of selective effects across the genome.

@@

Given these the sensitivity of mutation rate estimates and the strong-selection
non-identifiability issue, we next sought to assess the estimation accuracy of
our composite likelihood method using forward simulations. In order to have
enough megabase-scale data to apply our method, we independently evolved the
first five human chromosomes. Due to the immense computational expense of
conducting forward simulations at this scale over numerous parameter
combinations, we used fixed mutation and selection parameters in putatively
conserved regions (see Methods XXX). We note three important findings from
these simulations.

First, the linked selection signal, as measured by the $R^2$ between predicted
and simulated megabase-scale diversity, increases with the intensity of
selection against new deleterious mutations and mutation rate (Figure
\ref{fig:figure-4}B, top row). This is consistent across both classic BGS and
our B' method; the only exception is the classic BGS model run with a
full-coverage track for the reasons described above. However, if new mutations
are highly deleterious ($s=0.05$), $R^2$ is reduced. This is consistent with
very strong selection having less localized effects and acting more like a
genome-wide faster rate of drift.

Second, both classic BGS theory and our B' method accurately infer the average
selection coefficient under strong selection (Figure \ref{fig:figure-4}B,
middle row). However, when selection was weak, the classic BGS model
erroneously estimated strong selection and a very low mutation rate. By
contrast, our B' method estimated selection coefficients much closer to their
true value. A minor discrepancy occurs around $2Ns = 1$, for reasons discussed
in Section XXX. Because we only simulated fixed selection coefficients and five
chromosomes, we only assessed the accuracy of average selection coefficients
and not the DFE probabilities.

Third, with both our B' and the classic BGS models, we find mutation rate
estimates show slight biases in the strong-selection regime (Figure
\ref{fig:figure-3}B, bottom row). Still, our B' method leads to more accurate
mutation rate estimates than classic BGS theory across a variety of selection
coefficients. When BGS is not acting, either due to weak selection or a low
rate of deleterious mutations (Figure \ref{fig:figure-4}B right column), all
estimates are worse. However, we note that this is an unlikely region of
parameter space and this issue is readily diagnosable from the poor model fit 
as evident by low $R^2$ values.

Finally, we also used simulations to test the sensitivity of our method to
violations of two assumptions: constant population size and recessivity of
selective effects. Since we find little difference in parameter estimates among
YRI and bottlenecked out-of-Africa (CEU and CHB) samples, we tested whether a
population expansion would bias our method. We expansion rates like those
estimated for Europeans by \textcite{Gutenkunst2009-pg} had little effect on
parameter estimates (Supplementary Section \ref{supp:sim-assum}). By contrast,
we find our estimates are downwardly biased when mutational effects are weak
and recessive, but are accurate under recessive strong selection. This occurs
since the fixation probabilities of recessive mutations are in reality much
lower than they are under our model assuming additivity.

@@

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/figure_5.pdf} 
    \caption{}
  \label{fig:figure-5}
\end{figure}

While we have previously shown that out-sample predicted diversity matches
observed diversity (Section XXX), we also investigated whether there were
regions of the genome that were particularly poor-fitting. Comparing predicted
against observed diversity along chromosomes (see Supplementary Materials
Section \ref{supp:chrom-fit}) we overall find close correspondence consistent
with the high $R_\text{LOCO}^2$. However, we find a few large (tens of
megabases) regions with systematically poorer fit (Supplementary Materials
Figure \ref{supp:spatial-residuals}). In Figure \ref{fig:figure-5}A we see one
such region from XXX to XXX. Interestingly, predicted diversity closely follows
the peaks and troughs of this region. We note that a smaller segment within
this region has been found by a genome-wide scan for associative overdominance
\parencite{Gilbert2020-aw}.  For model XXX

>>>>

However, across all of our annotation models, new deleterious mutations in
conserved feature classes were consistently estimated to have strongly
deleterious effects (Figure \ref{fig:figure-3}A), consistent with previous work
\parencite{McVicker2009-ax,Murphy2022-sj}. This includes features such as
coding sequences, PhastCons regions, or the top 6\% of basepairs with the most
pathogenic CADD scores. However, we find that there is considerable uncertainty
in how strong selection is against the most deleterious mutations. For example,
in the CADD 6\% full-track model, new mutations are estimated to have a
selection coefficients of $s = 10^{-1}$ with 42\% chance, $s=10^{-2}$ with
7.8\% chance, and $s=10^{-3}$ with 41\% chance. This creates a bimodal DFE
(Figure \ref{fig:figure-3}A, second row, left column), which is mirrored across
the other CADD models. Moreover, these DFE estimates suggest a very high
average selection coefficient across CADD 6\% regions of $\widehat{\bar{s}}
\approx 0.043$ for YRI.

Given that previous methods have typically not estimated the DFE up to
selection coefficients of $s=10^{-1}$, we ran additional ``constrained grid"
models using a DFE only up to $s=10^{-2}$. Across all populations, these
constrained grid models fit the data less well than the default grid by about
one percentage point less in $R_\text{LOCO}^2$ (see Supplementary Table
\ref{supp:tbl-r2}). As expected, the constrained grid models distributed mass
onto the next strongest selection coefficient, $s=10^{-2}$, and the average
selection was reduced about six-fold to $\widehat{\bar{s}} \approx 0.0079$ for
YRI. While the DFE may truly be bimodal for highly-conserved regions and have a
very high average selection coefficient, a likely alternative is that these
results reflect a non-identifiability issue between selective effects of
$s=10^{-1}$ and $s=10^{-2}$. We will refer to this as the
\emph{strong-selection non-identifiability} issue.

@@

Finally, we note that mutation rate estimates in the Feature Priority model are
unreasonably high ($\widehat{\mu} \approx 4.2 \times 10^{-8}$), reminiscent of
the high mutation rate estimates found under McVicker et al.'s
(\citeyear{McVicker2009-ax}) model. While both our and Murphy et al.'s CADD and
PhastCons-based models alleviate this issue, it is worth discussing why this
could occur. We gain some insight from comparing the estimated mutation rates
between our Feature and PhastCons Priority annotation models, which both
contain the exact same number of basepairs, but these are assigned to different
feature classes based on the priority of overlapping features. That one of
these models is our best-fitting model, and the other our worst indicates the
sensitivity these models have when features classes have heterogeneous DFEs.
CADD-based models fit better in part due to their fine-scale resolution of
selective effects across the genome. While ideally we would fit a CADD model
with different features corresponding to the different percentiles of
pathogenicity, these features are on the basepair scale and thus too
memory-intensive for our method to currently accommodate.

@@

Given these the sensitivity of mutation rate estimates and the strong-selection
non-identifiability issue, we next sought to assess the estimation overall
accuracy of our composite likelihood method using forward simulations. In order
to have enough megabase-scale data to apply our method, we independently
evolved the first five human chromosomes and combined them into a ``synthetic"
genome. Due to the immense computational expense of conducting forward
simulations at this scale over numerous parameter combinations, we used fixed
mutation and selection parameters in putatively conserved regions (see Methods
XXX). We note three important findings from these simulations.

First, the linked selection signal, as measured by the $R^2$ between predicted
and simulated megabase-scale diversity, increases with the intensity of
selection against new deleterious mutations and mutation rate (Figure
\ref{fig:figure-4}B, top row). This shows the ``strength" of the linked
selection signal varies as expected with these key parameters. However, if new
mutations are highly deleterious ($s=0.05$), $R^2$ is reduced. This is
consistent with very strong selection having less localized effects and acting
more like a genome-wide faster rate of drift.

Second, both classic BGS theory and our B' method accurately infer the average
selection coefficient under strong selection (Figure \ref{fig:figure-4}B,
middle row). However, when selection was weak, the classic BGS model
erroneously estimates strong selection and a very low mutation rate. By
contrast, our B' method estimated selection coefficients much closer to their
true value. A minor discrepancy occurs around $2Ns = 1$, for reasons discussed
in Section XXX. Because we only simulated fixed selection coefficients and five
chromosomes, we only assessed the accuracy of average selection coefficients
and not the DFE probabilities.

Third, with both our B' and the classic BGS models, we find mutation rate
estimates show slight biases in the strong-selection regime (Figure
\ref{fig:figure-3}B, bottom row). Still, our B' method leads to more accurate
mutation rate estimates than classic BGS theory across a variety of selection
coefficients. These biases in mutation rate estimates suggest that benchmarking
genome-wide models via their concordance with pedigree-based rate estimates may
not be suitable. When BGS is not acting, either due to weak selection or a low
rate of deleterious mutations (Figure \ref{fig:figure-4}B right column), all
estimates are worse. This makes intuitive sense; the overall linked selection
signal is weaker relative to drift-based noise. However, we note that this is
an unlikely region of parameter space and this issue is readily diagnosable
from the low $R^2$ values.

Finally, we also used simulations to test the sensitivity of our method to
violations of two assumptions: constant population size and recessivity of
selective effects. Since we find little difference in parameter estimates among
YRI and bottlenecked out-of-Africa (CEU and CHB) samples, we tested whether a
population expansion would bias our method. We use an expansion rate and timing
based on those estimated for Europeans by \textcite{Gutenkunst2009-pg}, finding
they had little effect on parameter estimates (Supplementary Section
\ref{supp:sim-assum}). By contrast, we find our estimates are downwardly biased
when mutational effects are weak and recessive, but are accurate under
recessive strong selection. This occurs because the fixation probabilities of
recessive mutations are in reality much lower than under our model that assumes
additivity.

@@

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/figure_5.pdf} 
    \caption{(A) CADD 6\% full-track }
  \label{fig:figure-5}
\end{figure}

While we have previously shown that out-sample predicted diversity matches
observed diversity (Section XXX), we also investigated whether there were
regions of the genome that were particularly poor-fitting. Comparing predicted
against observed diversity along chromosomes (see Supplementary Materials
Section \ref{supp:chrom-fit}) we overall find close correspondence consistent
with the high $R_\text{LOCO}^2$. However, we find a few large (tens of
megabases) regions with systematically poorer fit (Supplementary Materials
Figure \ref{supp:spatial-residuals}). In Figure \ref{fig:figure-5}A we see one
such region from XXX to XXX. Interestingly, predicted diversity closely follows
the peaks and troughs of this region. We note that a smaller segment within
this region has been found by a genome-wide scan for associative overdominance
\parencite{Gilbert2020-aw}.
